---
title: "Movie Dataset Study"
author: "Deepti Gupta"
date: "April 24, 2019"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

options(scipen = 999)   # to disable the values to appear in exponential format

```


### Libraries and Packages
```{r load_packages,  message=FALSE, warning=FALSE}
#install.packages(readxl)
#install.packages(dplyr)
#install.packages(mice)
#install.packages(Hmisc)
#install.packages(plotly)
#install.packages(ggplot2)
#install.packages(GGally)
#install.packages(scales)
#install.packages(reshape2)
#install.packages(caTools)
#install.packages(gmodels)

library(readxl)
library(dplyr)
library(mice)
library(Hmisc)
library(plotly)
library(ggplot2)
library(GGally)
library(scales)
library(reshape2)
library(caTools)
library(gmodels)

```

### Load the dataset
```{r load_dataset}
movie <- read_excel("C:/MyRWork/IncedoCaseStudy-MovieDS/movie_metadata.xls")
View(movie)

colnames(movie)
dim(movie)
summary(movie)  

head(movie)
```

### Data Cleansing
```{r data_cleansing}
#Check the duplicates

NROW(movie[duplicated(movie),])  # 45 duplicate rows !!
movie[duplicated(movie),]

movie_sub = (unique(movie))

#Next step is to check variable types as per the data - numeric, character, logical, factor
str(movie_sub)
# there are certain column which should be converted into factors for better analysis and we can do that while dealing with nulls

md.pattern(movie_sub, rotate.names = TRUE)

# Below are the columns with missing values and numeric
# num_critic_for_reviews 
# director_facebook_likes 
# title_year 
# content_rating     
# aspect_ratio 
# budget 
# gross
# country 
# actor_1_facebook_likes 
# actor_2_facebook_likes 
# actor_3_facebook_likes
# facenumber_in_poster 
# duration 
# color 
# num_user_for_reviews 

#movie_sub$facenumber_in_poster[is.na(movie_sub$facenumber_in_poster)] = mean(na.omit(movie_sub$facenumber_in_poster))

movie_sub$num_critic_for_reviews <- as.numeric(impute(movie_sub$num_critic_for_reviews, mean))
movie_sub$director_facebook_likes <- as.numeric(impute(movie_sub$director_facebook_likes, mean))
movie_sub$title_year <- as.numeric(impute(movie_sub$title_year, median))
movie_sub$aspect_ratio <- as.numeric(impute(movie_sub$aspect_ratio, mean))
movie_sub$budget <- as.numeric(impute(movie_sub$budget, mean))
movie_sub$gross <- as.numeric(impute(movie_sub$gross, mean))
movie_sub$actor_1_facebook_likes <- as.numeric(impute(movie_sub$actor_1_facebook_likes, mean))
movie_sub$actor_2_facebook_likes <- as.numeric(impute(movie_sub$actor_2_facebook_likes, mean))
movie_sub$actor_3_facebook_likes <- as.numeric(impute(movie_sub$actor_3_facebook_likes, mean))
movie_sub$facenumber_in_poster <- as.numeric(impute(movie_sub$facenumber_in_poster, mean))
movie_sub$duration <- as.numeric(impute(movie_sub$duration, mean))
movie_sub$num_user_for_reviews <- as.numeric(impute(movie_sub$num_user_for_reviews, mean))


glimpse(movie_sub)

movie_sub$color = as.factor(movie_sub$color)
movie_sub$language = as.factor(movie_sub$language)
movie_sub$content_rating = as.factor(movie_sub$content_rating)
movie_sub$country = as.factor(movie_sub$country)

summary(movie_sub)

# the data looks better now, let's start with Exploratory data analysis
```


### Exploratory Data Analysis Data Preparation
```{r eda_data_prep}
# The question is - What is the recipe to make a blockbuster, profitable movie? 
# Let's calculate the profit of the movie to classify the blockbuster movies

#movie_sub$classified_movie_score <- ifelse(movie_sub$movie_score<8,(ifelse(movie_sub$movie_score<4,"Bad","Average")),"Good")

movie_sub$profit <- movie_sub$gross - movie_sub$budget
movie_sub$profit_perc <- round(((movie_sub$profit/movie_sub$budget)* 100),2)

# referring the below post, making an assumption on the criteria of blockbuster
# https://www.quora.com/What-are-the-criteria-for-a-movie-hit-super-hit-blockbuster-and-flop

# movie-budget < 1 million and profit > 150, blockbuster
# 1 million < movie-budget < 50 million and profit_perc > 250, blockbuster
# movie-budget > 50 million and profit_perc > 450, blockbuster else Not a blockbuster
 


movie_sub$is_blockbuster <- ifelse(movie_sub$budget< 1000000
                                   ,ifelse(movie_sub$profit_perc>150,1,0)      #budget < 1 million = yes
                                   ,ifelse(movie_sub$budget < 50000000       
                                                ,ifelse(movie_sub$profit_perc>250,1,0) #budget < 50 million = yes
                                                ,ifelse(movie_sub$profit_perc>450,1,0) #budget > 50 million 
                                          )
                                   )

table(movie_sub$is_blockbuster)
summary(movie_sub$profit)


ggcorr(movie_sub,label = TRUE,label_alpha = TRUE, hjust = 0.75, size = 2)

```

### Exploratory Data Analysis
```{r eda}
movie_sub$is_blockbuster = as.factor(movie_sub$is_blockbuster)
colnames(movie_sub)

# The years those produced the blockbuster movies
ggplot(movie_sub, aes(x=title_year, y=movie_score, color=is_blockbuster))+
  geom_jitter(alpha=0.5) +
  theme_bw() +   labs(title = "Blockbuster movies and their ImDB ratings over the years\n", 
       x = "Year of Release", y = "ImDB Movie Score", color = "Is Blockbuster?\n") +
  scale_color_manual(labels = c("No", "Yes"), values = c( "orange", "blue")) 
# Movies from recent years are more blockbusters. 
# Only blockbusters movies released before 1980s are listed for ImDB rating



# Question - Trends of popular movies on the basis of their duration
ggplot(data = movie_sub,aes(x=duration, fill = is_blockbuster)) +
  geom_histogram(stat = "count", binwidth = 10) + 
  labs(title = "Trends of duration of movies \n", 
       x = "Duration of the movie", y = "Length of Movie in Minutes", color = "Is Blockbuster?\n") +
    xlim(0, 200) + ylim(0, 100) +
    facet_wrap(~ is_blockbuster)
# Shorter the movie, more is the probability of it being popular



#Question - What is the overall grade distribution of the students?
CountryofMovieData <- as.data.frame(sort(table(movie_sub$country, dnn = c('Country')), decreasing = T),responseName = 'CountryofMovie')
CountryofMovieDataTop10 = CountryofMovieData[1:10,]

total = sum(CountryofMovieData$CountryofMovie)
ggplot(aes(x= reorder(Country,CountryofMovie), y = (CountryofMovie/total), fill = Country), data = CountryofMovieDataTop10) +
  geom_bar(stat = 'identity') +
  geom_text(stat='identity', aes(label = percent((CountryofMovie)/total)),
           data = CountryofMovieDataTop10, hjust = 0.5,vjust=-0.5) +
  theme(axis.text = element_text(size = 9)) +
  xlab("Top 10 Countries Classified") +
  ylab("Percentage of Movies Produced") + 
  ggtitle("Percentage of movies classified categories")
# USA Produced maximum number of movies


# Spread of critic's reviews - right skewed
ggplot(movie, aes(x = num_critic_for_reviews)) + geom_histogram(bins = 50, color="white", fill=rgb(0.2,0.7,0.1,0.4))  + ggtitle('Histogram of Critic Reviews')



# Trend of viewer's number of critics over the groups of years
movie_sub$year_groups = cut(movie_sub$title_year,breaks=c(1916,1950,1980,2000,2016), labels =c("1916-1950","1950-1980","1980-2000","2000-2016"))

ggplot(movie_sub) + 
  geom_density(aes(x = num_critic_for_reviews, color = year_groups, fill=year_groups), alpha=0.3) + 
  scale_colour_brewer(palette = 'Dark2') +
  facet_wrap(~is_blockbuster) 
# Non-block buster movies - Older movies are reviews more by the critics than the recent ones
# recent block buster movies are reviewed more than the recent non-block buster movies 


# The spread of movie ratings across years
# boxplot(movie_score ~ year_groups, data=movie_sub, col="seagreen") +
#  title("Movie Ratings Across the Years")
# Most of the movies were rated after 1980s


# After exploration, what could be the possible set of columns to influence a huge profit from a movie
str(movie_sub)
colnames(movie_sub)

# num_critic_for_reviews(3)
# duration  (4)
# num_voted_users (13)
# budget (23)
# movie_score (26)
# num_user_for_reviews (19)
# langauge (non-numeric)
ggpairs(movie_sub,columns=c(3,4,13,19,23,26))

# guess we need to remove budget as it doesn't show a significant influence
```

### Data Modelling
```{r data_model}
table(movie_sub$is_blockbuster)
1091/(1091+3907) # 21.8% movies are block buster

# Randomly split data
set.seed(54321)
split = sample.split(movie_sub$is_blockbuster, SplitRatio = 0.75)

movie_train = subset(movie_sub, split == TRUE)
movie_test = subset(movie_sub, split == FALSE)
nrow(movie_train)
nrow(movie_test)

logit_movie_1 = glm(is_blockbuster ~ num_critic_for_reviews + duration + num_voted_users  + 
                    budget  + movie_score  + num_user_for_reviews ,data=movie_train, family=binomial)
summary(logit_movie_1)
# Removing duration and num_user_for_reviews, as movie_Score is just on line, let's see if removing these variables
# have any impact on it

logit_movie_2 = glm(is_blockbuster ~ num_critic_for_reviews  + num_voted_users  + 
                    budget  + movie_score ,data=movie_train, family=binomial)
summary(logit_movie_2)
# Removing movie_Score as it is not significant


logit_movie_3 = glm(is_blockbuster ~ num_critic_for_reviews + num_voted_users + budget ,data=movie_train, family=binomial)
summary(logit_movie_3)

# Interpretation

# 1- Each one-unit change in num_critic_for_reviews will increase, the log odds of movie to be blockbuster decreases by 0.002, and its p-value indicates that it is somewhat significant in determining the hit movie. Same applicable for budget

# 2- Each one-unit change in num_voted_users will increase, the log odds of movie to be blockbuster increases by 0.000007, and its p-value indicates that it is somewhat significant in determining the hit movie.

# 3. The difference between Null deviance and Residual deviance tells us that the model is a good fit. Greater the difference better the model.

```

### Model Evaluation
```{r model_evaluation}
predict = predict(logit_movie_3, type = "response", newdata = movie_test)
table(movie_test$is_blockbuster, predict > 0.6)

CrossTable(x = movie_test$is_blockbuster, y = predict > 0.6, prop.chisq = FALSE)


# Accuracy

(1038)/(1250)   # 83% accuracy


```


SUMMARY

1. From EDA, movie's duration should be shorter
2. From model. number of critic's reviews should be lesser
3. Budget should not be too high
4. Votes should be very high

Congratulations! you have earned yourself a blockbuster